---
title: "AirBnb Dublin"
output: 
  html_notebook: 
    highlight: kate
---
Install the required packages and load the Original Dataset.
```{r}
install.packages(c("tidyverse","naniar","DataExplorer","ggpubr","lubridate","mice"))

library(naniar)
library(tidyverse)
library(dplyr)
library(DataExplorer)
library("ggpubr")
library(lubridate)
library(mice)

```
Read in the Rawdata
```{r}
rawdata <- read.csv("listings.csv")
```
Take a birdsEye overview of the raw dataset
```{r}
glimpse(rawdata)
```
Extract the relevant Variables to the study.
```{r}
DForiginal<- select(rawdata,id,price,neighbourhood_cleansed,market,latitude,longitude,property_type,room_type,accommodates,bathrooms,bedrooms,beds,bed_type,security_deposit,cleaning_fee,has_availability,number_of_reviews,review_scores_rating,cancellation_policy,calculated_host_listings_count,requires_license,host_id,host_since,host_is_superhost,host_total_listings_count,host_identity_verified)
```
Take a birdsEye overview of the selected  variables
```{r}
#glimpse(DForiginal)
```
Convert Variable Types 
```{r}
# AS Charactor
DForiginal$id <- as.character(DForiginal$id)
DForiginal$host_id <- as.character(DForiginal$host_id)
# AS Integers/Numeric
DForiginal$bathrooms <- as.integer(DForiginal$bathrooms)
DForiginal$security_deposit <- as.numeric(gsub("[\\$,]","",DForiginal$security_deposit))
DForiginal$price <- as.numeric(gsub("[\\$,]","",DForiginal$price))
DForiginal$cleaning_fee <- as.numeric(gsub("[\\$,]","",DForiginal$cleaning_fee))
# AS Date
DForiginal$host_since <- as_date(DForiginal$host_since)
```
MISSING VALUES & IMPUTATION 
```{r}
miss_var_table(DForiginal)
vis_miss(DForiginal)
miss_var_summary(DForiginal)

  # drop row for host_since as a Date can not be imputated by MICE package
which_na(DForiginal$host_since)
dforiginal <- DForiginal[-c(1172,6554), ]
```
```{r}

  # apply package MICE
init = mice(dforiginal, maxit=0) 
meth = init$method
predM = init$predictorMatrix
predM[, c("id","host_id")]=0
imputed = mice(dforiginal, method=meth, predictorMatrix=predM, m=1,seed=103)
  # view imputed values example
imputed$imp$review_scores_rating
  #Fill the NA with imputed values into new dataframe
df_imputed <- complete(imputed)
  # CONFIRMED NO MISSING VALUES IN imputed_df
miss_var_table(df_imputed)
miss_var_summary(df_imputed)
```
## HOT-ENCODE,BUCKETING CATAGORICAL FEATURES
```{r}
  # Identify all Catagorical variebles in dataframe
df_catagorical<-select_if(df_imputed,is.factor)
glimpse(df_catagorical)

  # First encode features with True|False creating a dummy variable of 1|zero
df_imputed <- df_imputed %>%
  mutate(hostverified_reg= ifelse(host_identity_verified=="t",1,0),
         superhost_reg=ifelse(host_is_superhost=="t",1,0),
         license_reg=ifelse(requires_license=="t",1,0),
         available_reg=ifelse(has_availability=="t",1,0))

df_imputed$hostverified_reg<-as.factor(df_imputed$hostverified_reg)
 # Second, bucketing other catagorical features together 


    # bucketting neighbourhood_cleansed
#table(df_imputed$neighbourhood_cleansed)
df_imputed <-  df_imputed %>%
  mutate(neighbourhood_cleansed= 
           case_when(neighbourhood_cleansed =="South Dublin" | neighbourhood_cleansed =="Dn Laoghaire-Rathdown" ~ "Out of City Center",
                     neighbourhood_cleansed =="Dublin City" | neighbourhood_cleansed =="Fingal" ~ "City Center"))

df_imputed$neighbourhood_cleansed<-as.factor(df_imputed$neighbourhood_cleansed)

    # Hot-Encode  neighbourhood_cleansed into saparate colmuns
df_imputed <- df_imputed %>%
  mutate(Incity_reg= ifelse(neighbourhood_cleansed=="City Center",1,0))
#glimpse(df_imputed$neighbourhood_cleansed)

    #bucketting Bed Types
#table(df_imputed$bed_type)
df_imputed <- df_imputed %>%
  mutate(bed_reg= ifelse(bed_type=="Real Bed",1,0))

    #bucketting Room Types
#table(df_imputed$room_type)
    # Entire home/apt = 1, (Private room + Shared room +Hotel room)=0
df_imputed <- df_imputed %>%
  mutate(FullApt_reg= ifelse(room_type=="Entire home/apt",1,0))


    # bucketting Cancellation Policy
table(df_imputed$cancellation_policy)
    # (flexible + moderate)= 1 ,(strict_14_with_grace_period + super_strict_30 + super_strict_60)= 0
df_imputed <- df_imputed %>%
  mutate(cancellation_reg= ifelse(cancellation_policy=="flexible"| cancellation_policy=="moderate",1,0))


    # bucketting Property Type
    # 33 types grouped into 3 ( House,Apt.Hotel,Other)
levels(df_imputed$property_type)

df_imputed <-  df_imputed %>%
  mutate(property_type= 
           case_when(property_type %in% c("House","Bungalow","Cabin","Chalet","Cottage","Guesthouse","House","Tiny house","Villa","Townhouse") ~ "House",
                     property_type %in% c("Hotel","Apartment", "Aparthotel", "Bed and breakfast","Boutique hotel","Condominium","Guest suite","Hostel","Loft","Serviced apartment") ~ "Apt.Hotel",
                     property_type %in% c( "Barn","Boat","Casa particular (Cuba)","Castle","Farm stay","Earth house","Houseboat","Hut","Igloo","Lighthouse","Nature lodge","Tent","Tipi","Other" ) ~ "Other_propertytype"))

df_imputed$property_type<-as.factor(df_imputed$property_type)


    # Look at a table of the new column created
table(df_imputed$property_type)
    # Hot-Encode  property_type into saparate colmuns
df_imputed <- df_imputed %>%
  mutate(House_reg= ifelse(property_type=="House",1,0))
df_imputed <- df_imputed %>%
  mutate(Apt.Hotel_reg= ifelse(property_type=="Apt.Hotel",1,0))
df_imputed <- df_imputed %>%
  mutate(Other_propertytype_reg= ifelse(property_type=="Other_propertytype",1,0))

```
Histograms
```{r}
# all numeric variables in df_imputed
df_numericencodeing <- data.frame(df_imputed$host_total_listings_count,df_imputed$calculated_host_listings_count,df_imputed$review_scores_rating,df_imputed$number_of_reviews,df_imputed$cleaning_fee,df_imputed$security_deposit,df_imputed$beds,df_imputed$bedrooms,df_imputed$bathrooms,df_imputed$accommodates,df_imputed$price)

```

```{r}
hist(log(df_imputed$price),xlim = c(0,20), col = "green")
hist((df_imputed$calculated_host_listings_count), col = "grey")
hist((df_imputed$review_scores_rating) , col = "grey")
hist((df_imputed$number_of_reviews), col = "grey")
hist((df_imputed$cleaning_fee), col = "grey")
hist((df_imputed$security_deposit), col = "grey")
hist((df_imputed$beds), col = "grey")
hist((df_imputed$bedrooms), col = "grey")
hist((df_imputed$bathrooms), col = "grey")
hist((df_imputed$accommodates), col = "grey")
hist((df_imputed$host_total_listings_count), col = "grey")

```
EDA
 Create contingency table, which is a useful way to represent the total counts of observations
```{r}
#table(df_imputed$cleaning_fee)
#table(df_imputed$calculated_host_listings_count)
#table(df_imputed$review_scores_rating)
#table(df_imputed$number_of_reviews)
#table(df_imputed$security_deposit)
#table(df_imputed$beds)
#table(df_imputed$bedrooms)
#table(df_imputed$bathrooms)
#table(df_imputed$accommodates)
#table(df_imputed$host_total_listings_count)

```



```{r}
# extract the year from host_since date
host_years <- year(df_imputed$host_since)
head(host_years)
class(host_years)
table(host_years)
plot_histogram(host_years)
```
MODEL CONSTRUCTION (REGGRESSION)
```{r}
# normalise the distribustion of listtnings Prises - DEPENDENT 
price <- df_imputed$price
price <- as.numeric(gsub("[\\$,]","",df_imputed$price))
glimpse(price) # overview
which_na(price) # identify any missing prices
hist(price) # graph distribution
log_price<-log(price) # graph distribution of logarithim of price
hist(log_price)
```
# STEPWISE REGRESSION using the numeric features ONLY.
```{r}
# select the numeric variables and store in a dataframe
df_numericencodeing <- data.frame(df_imputed$host_total_listings_count,df_imputed$calculated_host_listings_count,df_imputed$review_scores_rating,df_imputed$number_of_reviews,df_imputed$cleaning_fee,df_imputed$security_deposit,df_imputed$beds,df_imputed$bedrooms,df_imputed$bathrooms,df_imputed$accommodates)

# Regression model_1
fitstart<-lm(df_imputed$price~1,data=df_numericencodeing)
fitall=lm(df_imputed$price ~.,df_numericencodeing)
formula(fitall)

# Stepwise Forward & Backwords
model_1<-step(fitstart,direction="both",scope = formula(fitall))
summary(model_1)
```
# STEPWISE REGRESSION using numeric and encoded catagorical features.
```{r}
# select the numeric and incoded catagorical variables and store in a dataframe
regdata<-df_imputed[ ,c(9,10,11,12,14,15,17,18,20,25,27:37)]

# Regression model_2
fitstart2<-lm(df_imputed$price~1,data=regdata)
fitall2=lm(df_imputed$price ~.,regdata)
formula(fitall2)

# Stepwise Forward & Backwords
model_2<-step(fitstart2,direction="both",scope = formula(fitall2))
summary(model_2)
```


EDA examples
Side-by-side box plots
Faceted histograms
Overlaid density plots
```{r}
df_imputed %>% ggplot(aes( y = log_price,color=property_type)) +
 stat_boxplot()
```
```{r}
df_imputed %>% ggplot(aes(log_price+0.10,fill=neighbourhood_cleansed)) +
 geom_histogram(bins=50)
```
```{r}
df_imputed %>% ggplot(aes( y = log_price,fill=room_type)) +
 stat_boxplot()
```
```{r}

df_imputed %>% ggplot(aes(x=df_imputed$hostverified_reg,fill=hostverified_reg)) +
 geom_bar()
df_imputed %>% ggplot(aes(x=df_imputed$bedrooms)) +
 geom_bar()
df_imputed %>% ggplot(aes(x=df_imputed$superhost_reg,fill=hostverified_reg)) +
 geom_bar()
df_imputed %>% ggplot(aes(x=df_imputed$cancellation_reg)) +
 geom_bar()
df_imputed %>% ggplot(aes(x=df_imputed$host_total_listings_count,fill=df_imputed$Incity_reg)) +geom_bar()
```

```{r}

df_imputed %>% ggplot(aes(x=host_total_listings_count,y=log_price,fill=property_type)) +
 geom_boxplot()
```

```{r}
table(df_imputed$cancellation_policy)
df_imputed %>% group_by(year(host_since))%>%
  ggplot(aes(y=price,x=host_since))+geom_point()
```

```{r}
install.packages("corrplot")
library(corrplot)
```
```{r}
names(df_numericencodeing)<- c("review score","review number","total listings","cal_list_count","deposit","cleaning fee","bathrooms","bedrooms","beds","total_people")
m<-cor(df_numericencodeing)
corrplot(m, method="color",type="upper",order="hclust")

```
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(M, method="color", col=col(200),  
         type="upper", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         p.mat = p.mat, sig.level = 0.01, insig = "blank", 
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
         )
         
```{r}
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(m, method="color", col=col(200),  
         type="upper", order="hclust", 
         addCoef.col = "grey", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # hide correlation coefficient on the principal diagonal
         diag=F
         )
```

```{r}
o<- ggplot(data = df_imputed, aes(x=df_imputed$property_type, y=df_imputed$cancellation_policy, color=cancellation_policy))
 o + geom_jitter(alpha=0.3) + scale_color_manual(breaks = c('Apt.Hotel','House','Other_propertytype'),values=c('navyblue','pink','darkgreen','purple','red'))
 
 p<- ggplot(data = df_imputed, aes(x=df_imputed$property_type, y=df_imputed$neighbourhood_cleansed, color=neighbourhood_cleansed))
 p + geom_jitter(alpha=0.3) + scale_color_manual(breaks = c('Apt.Hotel','House','Other_propertytype'),values=c('Turquoise','darkorange'))
```
Refernce"
http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/78-perfect-scatter-plots-with-correlation-and-marginal-histograms

https://www.r-bloggers.com/to-eat-or-not-to-eat-thats-the-question-measuring-the-association-between-categorical-variables/
